# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
daily_steps <- aggregate(data$steps, by=list(data$date), FUN=sum, na.rm=TRUE)
names(daily_steps) <- c("day", "steps")
hist(daily_steps$steps, xlab = "Steps", main = "Daily Steps")
```

```{r}
steps_mean <- as.integer(round(mean(daily_steps$steps, na.rm = TRUE), 0))
steps_median <- median(daily_steps$steps, na.rm = TRUE)
```

Daily step mean count is `r steps_mean`.  
Daily step median count is `r steps_median`.

## What is the average daily activity pattern?

```{r}
interval_steps <- aggregate(data$steps, by=list(data$interval), FUN="mean", na.action=na.pass, na.rm=TRUE)
names(interval_steps) <- c("interval", "steps")
plot(interval_steps, type = 'l', xlab = 'Interval', ylab = 'Average Steps')
max_steps <- max(interval_steps$steps)
max_interval <- interval_steps$interval[interval_steps$steps == max_steps]
```

The interval that contains the average maximum number of steps is `r max_interval`.

## Imputing missing values

```{r}
na_values <- sum(is.na(data$steps))
```

Data set contains `r na_values` NA values. Na values are replaces by interval's mean value.

```{r}
new_data <- data
for(i in 1 : nrow(new_data)) {
    if(is.na(new_data$steps[i])) {
        new_data$steps[i] <- as.integer(round(interval_steps$steps[interval_steps$interval == new_data$interval[i]], 0))
    }
}

new_daily_steps <- aggregate(new_data$steps, by=list(new_data$date), FUN=sum)
names(new_daily_steps) <- c("day", "steps")
hist(new_daily_steps$steps, xlab = "Steps", main = "Daily Steps")
```

```{r}
new_steps_mean <- as.integer(round(mean(new_daily_steps$steps), 0))
new_steps_median <- median(new_daily_steps$steps)
```

New daily step mean count is `r new_steps_mean`. New daily step median count is `r new_steps_median`.  
Mean and median values are higher than the earlier estimates.  This is because data with NA values had more days with 0 steps.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekend <- c('lauantai', 'sunnuntai') # Saturday and Sunday in Finnish
new_data$weekend[weekdays(as.Date(new_data$date)) %in% weekend] <- 'weekend'
new_data$weekend[is.na(new_data$weekend)] <- 'weekday'
new_data$weekend <- as.factor(new_data$weekend)

weekend_steps <- aggregate(new_data$steps, by=list(new_data$interval, new_data$weekend), FUN=mean)
colnames(weekend_steps) <- c("interval", "weekend", "steps")
library(ggplot2)
ggplot(weekend_steps, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .) +
    xlab("Interval") + ylab("Steps")
```

During weekdays there is an activity spike in the morning. During weekends activitys are more evenly distributed, there are multiple high points circa 150 steps during the day.